Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-14 shanghai.tex
All mention of picard groups removed, except for the introduction.

  The sequence of Picard groups is right-exact: it is exact at
  $\QCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\QCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker t_{\pi_0(G)} =0$
  (by Propositions~\ref{prop:etale} and \ref{prop:X}) and
  \[
  \image(\QCiso{\pi_0(G)} \to \QCiso{G}) \subseteq \ker(\QCiso{G} \to \QCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\cs{L}$ is a quasicharacter sheaf on $G$ that is trivial on
  $G^0$. Then $\trFrob{\cs{L}}\vert_{G^0(\Fq)} = \trFrob{\cs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\cs{L}}$. By Propositions~\ref{prop:etale} and \ref{prop:X},
  there is a a quasicharacter sheaf $\cs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\cs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\cs{L}$ is isomorphic to the
  image of $\cs{L}_\chi$ under the monoidal functor $\QC(\pi_0(G)) \to \QC(G)$.

@@

\section{Greenberg of Neron}
\label{sec:GN}

%\subsection{The N\'eron locally finite type model for $T$}

@@

%\input{application}
\begin{theorem}\label{thm:application}
For each positive integer $d$, the trace of Frobenius 
  \[
t_{T,d}:  \QCiso{\GN{T}_d} \to \Hom_{< d}(T(K),\EEx)
  \]
is an isomomorphism from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{<d}(T(K),\EEx).
\]
\end{theorem}

\begin{proof}
  By \ref{GNd}, we may apply Theorem~\ref{thm:snake} to
  $\GN{T}_d$ which provides a canonical isomorphism
  $\QCiso{\GN{T}_d} \iso \GN{T}_d(\Fq)^*$.
  By \ref{dFq}, $\GN{T}_d(\Fq) = \TT(A_d)$.
  To finish the proof we recall that the filtration of $T(K)$ used to
  define depth has the property that $\TT(A_d) \cong T(K) / T(K)_d$
  \cite[Prop 5.2]{Yu}, so characters of $\GN{T}_d(\Fq)$ 
  are in bijection with characters of $T(K)$ vanishing on $T(K)_d$.

  What do we mean by saying
\end{proof}

%\section{Application to quasicharacters}

@@

This gives us a natural way to geometrize all quasicharacters of $T(K)$ without reference to depth.

\begin{corollary}
Let $T$ be an algebraic torus over a non-Archimedean local field $K$.
Then
  \[
 \QCiso{\GN{T}} \iso \Hom_\text{ad}(T(K),\EEx),
  \]
  canonically.
\end{corollary}

\section{Functoriality}

\begin{proposition}

\end{proposition}

\begin{proof}
When combined with the definition of $\QC(\GN{T})$ and Property~\ref{T(K)}, the first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor, while the second part shows that Trace of Frobenius is a natural transformation $t: F_1 \to F_2$. When further combined with Lemma~\ref{prop:product}, we see that $F_1$ is an additive functor and $t: F_1 \to F_2$ is a natural transformation between additive functors.
\end{proof}

\section{Geometrization of restriction of scalars of local fields}

@@

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k} \Gr(\TL)} \arrow[two heads]{r} \dar & \QCiso{\GN{T}} \dar\\
\Hom(T(L), \EEx) \arrow[two heads]{r} & \Hom(T(K), \EEx)
\end{tikzcd}
\]
\end{proposition}

>>>>

  The sequence of groups of isomorphism classes of quasicharacter sheaves is right-exact: it is exact at
  $\QCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\QCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker t_{\pi_0(G)} =0$
  (by Propositions~\ref{prop:etale} and \ref{prop:X}) and
  \[
  \image(\QCiso{\pi_0(G)} \to \QCiso{G}) \subseteq \ker(\QCiso{G} \to \QCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\cs{L}$ is a quasicharacter sheaf on $G$ that is trivial on
  $G^0$. Then $\trFrob{\cs{L}}\vert_{G^0(\Fq)} = \trFrob{\cs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\cs{L}}$. By Propositions~\ref{prop:etale} and \ref{prop:X},
  there is a a quasicharacter sheaf $\cs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\cs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\cs{L}$ is isomorphic to the
  image of $\cs{L}_\chi$ under the monoidal functor $\QC(\pi_0(G)) \to \QC(G)$.

@@

\section{Greenberg of Neron}
\label{sec:GAN}

@@

\section{Application to quasicharacters of bounded depth}

\begin{theorem}\label{thm:application}
Let $d$ be a positive integer.
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T,d}:  \QCiso{\GN{T}_d} \to \Hom_{< d}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T,d}$ is functorial in the following sense: $T \mapsto \trFrob{T,d}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}_d}\]
to the additive functor
\[
T \mapsto \Hom_{<d}(T(K),\EEx).
\]
\item
For every finite Galois extension $L/K$ and for every algebraic torus $T$ over $K$, the following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Gr(\underline{\Res_{L/K}(T_L)}_d)} \arrow[two heads]{r} \dar & \QCiso{\Gr({T}_d)} \dar\\
\Hom_{<d}(T(L), \EEx) \arrow[two heads]{r} & \Hom_{<d}(T(K), \EEx)
\end{tikzcd}
\]
\end{enumerate}
\end{theorem}

\begin{proof}
  By \ref{GNd}, we may apply Theorem~\ref{thm:snake} to
  $\GN{T}_d$ which provides a canonical isomorphism
  $\QCiso{\GN{T}_d} \iso \GN{T}_d(\Fq)^*$.
  By \ref{dFq}, $\GN{T}_d(\Fq) = \TT(A_d)$.
  Then we recall that the filtration of $T(K)$ used to
  define depth has the property that $\TT(A_d) \cong T(K) / T(K)_d$
  \cite[Prop 5.2]{Yu}, so characters of $\GN{T}_d(\Fq)$ 
  are in bijection with characters of $T(K)$ vanishing on $T(K)_d$. 
  This shows that $\trFrob{T,d}$ is an isomorphism.

  Next we show that this isomorphism is functorial.
  When combined with the definition of $\QC(\GN{T})$ and Property~\ref{T(K)}, the first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor, while the second part shows that Trace of Frobenius is a natural transformation $t: F_1 \to F_2$. When further combined with Lemma~\ref{prop:product}, we see that $F_1$ is an additive functor and $t: F_1 \to F_2$ is a natural transformation between additive functors.

  Finally, we turn to restriction of scalars.
\end{proof}

\section{The geometrization of quasicharacters}

@@

\begin{corollary}\label{cor:application}
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T}:  \QCiso{\GN{T}} \to \Hom_{\text{ad}}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T}$ is functorial in the following sense: $T \mapsto \trFrob{T}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{\text{ad}}(T(K),\EEx).
\]
\item
For every finite Galois extension $L/K$ and for every algebraic torus $T$ over $K$, the following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\underline{\Res_{L/K}(T_L)}} \arrow[two heads]{r} \dar & \QCiso{\GN{T}} \dar\\
\Hom_{\text{ad}}(T(L), \EEx) \arrow[two heads]{r} & \Hom_{\text{ad}}(T(K), \EEx)
\end{tikzcd}
\]
\end{enumerate}
\end{corollary}

@@
